What We Eval in the Shadows

### 2021-04-15 paper/main.tex
Example for F(F(X))

A peculiar use of \eval is to bypass some checks performed to accept code submitted to CRAN. For example, bindings can be \emph{locked} in an environment, \ie forbidding its value to be modified. The binding can be unlocked though by calling \c{unlockBinding} but packages in CRAN are forbidden to use that function.  Package \emph{data.table} bypasses this restriction by building a call piece by piece, to unlock binding \c{.N}. The call is then evaluated. Now that \c{.N} is unlocked, it can be assigned to. Instead of using \eval to assign in the parent frame, they rather use the more specialized function \c{assign}.

@@

>>>>

A peculiar use of \eval is to bypass some checks performed to accept code submitted to CRAN. For example, bindings can be \emph{locked} in an environment, \ie forbidding its value to be modified. The binding can be unlocked though by calling \c{unlockBinding} but packages in CRAN are forbidden to use that function.  Package \emph{data.table} bypasses this restriction by building a call piece by piece, to unlock binding \c{.N}. The call is then evaluated. Now that \c{.N} is unlocked, it can be assigned to. Instead of using \eval to assign in the parent frame, they rather use the more specialized \c{assign} function.

@@

\mypara{Complex function calls.} It corresponds to minimized expressions \c{F(X)} and \c{F(F(X))}, \ie function calls with variables, possibly with nested function calls. Package \emph{huxtable} is used to create custom tables and export them to various formats. To create functions to customize each of the borders of the table, it injects the indices \c{i} and \c{j} of the cells at the border and then evaluate them:
\begin{lstlisting}
xxx_border_huxtable <- function(lr_tb, i, j, border_prop) {
	function (ht) {
		hickness <- mlist$thickness[eval(i), eval(j), drop = FALSE]
		...
	}
}
left_border.huxtable <- xxx_border_huxtable("lr_borders",
i = quote(seq_len(nrow(ht))), j = quote(-last_col))
\end{lstlisting}
